---
title: "Introduction to Data Visualization Using ggplot2 - Part 2"
output: html_document
author: OmaymaS
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width = 10)
```

In the following exercises you will explore the [googleplaystore](https://raw.githubusercontent.com/OmaymaS/datafest_tidyverse_workshop/master/data/googleplaystore.csv) dataset; a slightly modified version of the original [Google Play Store Apps](https://www.kaggle.com/lava18/google-play-store-apps#googleplaystore.csv) on Kaggle.  You will answer questions related to:

- applications reviews per category.
- free versus paid applications per category.
- the distribution of application size over different categories.

# Load Libraries 

```{r load libraries}
## load libraries 
library(ggplot2)  ## for data visualization
library(dplyr)    ## for data manipulation
library(readr)    ## to use read_csv

## set default theme
theme_set(theme_minimal())
```

# Import Data

```{r read data}
## read data from github
googleplaystore <- read_csv("https://raw.githubusercontent.com/OmaymaS/datafest_tidyverse_workshop/master/data/googleplaystore.csv")
```

```{r explore googleplaystore}
## explore googleplaystore
googleplaystore
```

# Explore Data

## 1. Applications reviews per `category`

### What is the total number of reviews for each category?
(same with top 10 in terms of apps)

#### 1.1

- calculate the total number of reviews per category `reviews_all`.
- arrange the dataframe (in a descending order) according to `reviews_all`.

```{r 1.1.1}
## calculate the number of reviews by category
apps_reviews <- googleplaystore %>% 
  group_by(## [WRITE CODE]) %>% 
  summarize(## [WRITE CODE]) %>% 
  arrange(## [WRITE CODE])

## view the resulting dataframe
apps_reviews
```

#### 1.2

- Extract the entries corresponding to the top ten categories in terms of total reviews.

```{r 1.2.1}
## select the first ten rows from apps_reviews_10
apps_reviews_10 <- ## [WRITE CODE]
```

- Create bar chart with `reviews_all` versus `category`. *Look at the order of the bars, what do you notice?*

```{r 1.2.2}
## create bar chart with reviews_all versus category
ggplot(## [WRITE CODE])+
  geom_bar(## [WRITE CODE])
```

#### 1.3

- Recreate the previous plot with the bars ordered in a descending order based on `reviews_all`.
- Assign the result to `top_10` and view it.

```{r 1.3}
## create a bar plot with a descending order based on reviews_all
top_10 <- ggplot(data = apps_reviews_10,
       aes(x = reorder(## [WRITE CODE]),
           y = reviews_all))+
  geom_bar(stat = "identity")

## view the plot
top_10
```

#### 1.4

- Modify `top_10` plot to flip the coordinates. 

```{r 1.4}
## flip the coordinates of top_10
top_10+
  ## [WRITE CODE]
```

*What do you notice about the order of the bars?*

## 2. Free Versus paid applications

### What is the number of free/paid applications in each category?

(same with teens)

#### 2.1

- Calculate the number of applications in each (`category`, `type`) group.
- Assign the result to `apps_types` and view it.

```{r 2.1}
## calculate the number of applications per (category, type) groups
 apps_types <- googleplaystore %>% 
  group_by(## [WRITE CODE]) %>% 
  summarize(apps = ## [WRITE CODE])

## view the result
apps_types
```

#### 2.2

- Create a bar plot with the number of apps versus each category.
- Color the bars by the application `type`. *Pick the right attribute to set!*
- Flip the coordinates to have horizontal bars.

```{r 2.2, fig.height=8}
## plot the number of applications vs. category, and color the bars by type
## [WRITE CODE]
```

*Can you spot a category with zero paid applications?*

### What is the share of free/paid applications in each category?

#### 2.3

- Recreate the previous plot with the percentage of applications instead of the count.

**HINT:**consider using the `position` attribute.

```{r 2.3, fig.height=8}
## recreate the previous plot with share instead of count
ggplot(data = apps_types, aes(x = category, y = apps, fill = type))+
  geom_bar(stat = "identity", ## [WRITE CODE])+
  coord_flip()
```

*Try `position = "dodge"`, what do you notice?*

## 3. Application size distribution

### What is the distribution of applications `size`?

#### 3.1 (Single category)

- Extract the entries corresponding to the FAMILY `category`
- Assign the result to `apps_family`

```{r 3.1.1}
## extract the entries corresponding to the FAMILY category
apps_family <- googleplaystore %>% 
  ## [WRITE CODE]
```

- Create a histogram to show the distribution of `size` in `apps_family`.

```{r 3.1.2}
## create a histogram of size in apps_family data
ggplot(## [WRITE CODE])+
  geom_histogram()
```

*Notice how the the distributions are skewed. The high values can be interesting for further exploration!*

#### 3.2 (Multiple categories)

- Extract the entries corresponding to the categories *("FAMILY","GAME", "TOOLS", "MEDICAL", "BUSINESS", "PRODUCTIVITY")*.
- Assign the result to `apps_top`

```{r 3.2.1}
## extract the data corresponding to the 6 given categories
apps_top <- googleplaystore %>% 
  filter(category %in% c("FAMILY","GAME", "TOOLS", "MEDICAL", "BUSINESS", "PRODUCTIVITY"))
```

- Plot the distribution of the application `size` in `apps_top`.
- Color the histograms by `category`.
- Use `facet_wrap()` to separate the six categories in six plots.

```{r 3.2.2}
## create a plot with 6 histograms corresponding to the 6 groups
ggplot(data = apps_top, aes(size, fill = category))+
  ## [WRITE CODE]
```

#### 3.3 (Multiple categories: a closer look!)

Look closer at the majority of applications:

- Recreate the previous Plot with the histograms.
- Set the `binwidth` to 5. 
- Set the x-axis limits to `(0, 150)`. 

```{r 3.3}
## modify the binwidth and x-axis limit
ggplot(data = apps_top, aes(size, fill = category))+
  geom_histogram(## [WRITE CODE])+
  coord_cartesian(## [WRITE CODE])+
  facet_wrap(~category)+
  labs(title = "Distribution of the size of android applications on googleplay",
       subtitle = "(Only applications with size <= 150 in the dataset)")
```

*What else could you find in this dataset? keep exploring!*