---
title: "Introduction to Data Visualization Using ggplot2"
output: html_document
author: OmaymaS
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
```

In the following set of exercises you will explore patterns and trends in the `gapminder` dataset.

**First**, you need to install and load the following libraries:

```{r load libraries}
library(ggplot2)    ## for data visualization
library(dplyr)      ## for data manipulation
library(gapminder)  ## to use gapmider dataset
```

Now, you are ready to follow the instructions and complete any missing code indicated by `## [WRITE CODE]`.

## Grammer of Graphics

Components:

- Data
- Aesthetics
- Geometries 
- Facets
- Statistics
- Coordinates 
- Themes 

### Data and Aesthetics

#### 1.1

- Explore the `gapminder` dataset.

**NOTE**: To see a detailed description of the variables, type `?gapminder`

```{r 1.1.1}
## show the gapminder dataframe
gapminder 
```

```{r 1.1.2}
## show a summary of the dataset
str(gapminder)
```

#### 1.2

- Extract the entries corresponding to year 2007 using `filter()`.
- Assign the resulting dataframe to `gapminder_2007`.

```{r 1.2}
## extract the entries of year 2007
gapminder_2007 <- gapminder %>% 
  filter(year == 2007)
```

#### 1.3

- Create a scatter plot with `lifeExp` versus `gdpPercap`.

```{r 1.3}
## plot `lifeExp` versus `gdpPercap`
ggplot(data = gapminder_2007, aes(x = gdpPercap, y = lifeExp))+
  geom_point()
```

#### 1.4

- Recreate the previous plot.
- Change the color of ALL the points to "blue".

```{r 1.4}
## recreate the previous plot and set the points color to blue
ggplot(data = gapminder_2007, aes(x = gdpPercap, y = lifeExp))+
  geom_point(color = "blue")
```

#### 1.5

- Recreate the previous plot.
- Color the points by `continent` instead of a single color.

```{r 1.5}
## recreate the previous plot and color the points by continent
ggplot(data = gapminder_2007, aes(x = gdpPercap, y = lifeExp,
                                  color = continent))+
  geom_point()
```

#### 1.6

- Recreate the previous plot
- Set the point size according to the population `pop`.

```{r 1.6}
## recreate the previous plot and set the size of points to pop
ggplot(data = gapminder_2007, aes(x = gdpPercap, y = lifeExp,
                                  color = continent,
                                  size =  pop))+
  geom_point()
```

*How would you give all the points a fixed size?*

#### 1.7

- Recreate the previous plot.
- Set the points transparency to 0.75 using the `alpha` attribute.

```{r 1.7}
## recreate the previous plot and set  alpha to 0.75
ggplot(data = gapminder_2007, aes(x = gdpPercap, y = lifeExp,
                                  color = continent,
                                  size =  pop))+
  geom_point(alpha = 0.75)
```

### Facets (Small Multiples)

**What if you wanted to have a separate plot for each continent on the same scale?** 

First we will assign the scatter plot to an object `plot_2007`. You can view the figure by typing `plot_2007`. And you can modify it by adding other layers.

```{r}
## create scatter plot with lifeExp Vs. gdpPercap and assign to plot_2007
plot_2007 <- ggplot(data = gapminder_2007, aes(x = gdpPercap, y = lifeExp,
                                               color = continent,
                                               size =  pop))+
  geom_point(alpha = 0.75)

## view plot_2007
plot_2007
```

#### 2.1

- Modify `plot_2007` using `facet_wrap()` to have a separate plot for each continent. *Notice how the factes are distrubuted in rows and columns*.

```{r 2.1, fig.width=10}
## modify plot_2007 to have a separate plot for each continent
plot_2007+
  facet_wrap(~continent)
```

#### 2.2

- Modify `plot_2007` to distribute the five plots in one row.

```{r 2.2, fig.width=12}
## organize the plots corresponding to the 5 continents in 1 row
plot_2007+
  facet_wrap(~continent, nrow = 1)
```

### Coordinates

**Now, what if you wanted to have a log scale on the x axis?**

#### 3.1

- Using `gapminder_2007`:

    + create a scatter plot with `lifeExp` versus `log(gdpPercap)`. 
    + color the points by continent.

```{r 3.1}
## plot lifeExp versus log(gdpPercap)
ggplot(data = gapminder_2007, aes(x = log(gdpPercap), y = lifeExp,
                                  color = continent))+
  geom_point()
```

#### 3.2

- Recreate the previous plot in a different way:

    + plot `lifeExp` Vs. `gdpPercap`.
    + use `scale_x_log10()` to change the x axis scale

```{r 3.2}
## plot lifeExp versus gdpPercap and use scale_x_log10()
ggplot(data = gapminder_2007, aes(x = gdpPercap, y = lifeExp,
                                  color = continent))+
  geom_point()+
  scale_x_log10()
```

**NOTE:** the x axis label could be misleading, because it is actually `log(gdpPercap)` not `gdpPercap`.

**What if you wanted to change the axes labels?**

#### 3.3

- change the x-axis label to `log(gdpPercap)` using `labs()`.

*What other labels and titles can you set using `labs()`?*

```{r 3.3}
## change the x-axis label to log(gdpPercap)
ggplot(data = gapminder_2007, aes(x = gdpPercap, y = lifeExp,
                                  color = continent))+
  geom_point()+
  scale_x_log10()+
  labs(x = "log(gdpPercap)")
```

### Themes 

**What if you wanted to change the gray background of the plots?**

#### 4.1

- Use `theme_minimal()` to change the theme of `plot_2007`. *Try to experiment with other themes!*

```{r4.1}
## use `theme_minimal()`
plot_2007+
  theme_minimal()
```

### Multiple Geoms/Layers

In the next set of exercises, you will explore and compare the change in `lifeExp` for four African countries.

#### 5.1

- Extract the entries corresponding to the four countries **["Rwanda", "South Africa", "Mozambique", "Kenya"]**.

```{r 5.1}
## extract the entries of the 4 countries
countries_africa <- gapminder %>% 
  filter(country %in% c("Rwanda", "South Africa", "Mozambique", "Kenya")) 
```

#### 5.2

- Create a line plot with `lifeExp` change over time for the countries in `countries_africa`.
- Set the line size to 1.3

*Can you see 4 lines corresponding to the 4 countries?*

```{r 5.2}
## plot lifeExp over time for the 4 countries
ggplot(data = countries_africa, aes(x = year, y = lifeExp,
                                    color = country))+
  geom_line(size = 1.3)
```

#### 5.3

- Create a line plot with `lifeExp` change over time for the countries in `countries_africa`.
- Add a points layer to the following line plot.
- Set the theme to `theme_minimal()`.

```{r 5.3}
## add a points layer
ggplot(data = countries_africa, aes(x = year, y = lifeExp,
                                    color = country))+
  geom_line()+
  geom_point()+ 
  theme_minimal()
```

## Interactivity with `ggplotly` (Extra)

**What if you wanted to see more details about each point?**

#### 6.1

- use `ggplotly()` function to turn `plot_2007` into an interactive plot.

*Hover over the points and check the tools on the upper toolbar.*

```{r 6.1}
## load library
library(plotly)

## pass plot_2007 to ggplotly
ggplotly(plot_2007)
```