---
title: "Getting Started with Data Manipulation Using dplyr"
output: html_document
author: OmaymaS
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Load libraries

```{r load libraries}
## load libraries 
library(readr) ## to use read_csv 
library(dplyr) ## for data manipulation
```

## Read data

```{r import data}
## read data from github
ge_trade <- read_csv("https://raw.githubusercontent.com/OmaymaS/datafest_tidyverse_workshop/master/data/ge_trade")
```

- Explore the `ge_trade` dataset.

```{r explore data}
## explore ge_trade
ge_trade
```

**NOTE:** The `imports_usd` and `exports_usd` values are in thousands.

## Manipulate data

### 1- The pipe` %>%`

#### 1.1

- Use the pipe to write an equivalent expression to `names(ge_trade)`.

```{r 1.1.1}
## rewrite names(ge_trade) using the pipe
## [WRITE CODE] ##
```

- rewrite `unique(ge_trade$region)` using the pipe.

```{r 1.1.2}
## rewrite unique(ge_trade$region) using the pipe
## [WRITE CODE] ##
```

**NOTE:** These are simple examples. However, the power of the pipe appears when you have multiple functions calls in series.

### 2- select()

#### 2.1- Select specific columns

- select `country`, `year`, `month` and `imports_usd`, `exports_usd`. *Think about different ways to achieve this!*
- select all the columns except for `code`.

```{r 2.1 method_1}
## method 1
ge_trade %>% 
  ## [WRITE CODE] ##
```

#### 2.2- Specify columns to exclude

- select all columns except `code` and `region`.

```{r 2.2}
## select all columns except code and region
## [WRITE CODE] ##
```

### 3- filter()

#### 3.1- Single condition

- Extract the entries corresponding to Australia.

```{r 3.1.1}
## extraxt the entries of Australia 
ge_trade %>% 
  ## [WRITE CODE] ##
```

- Extract the entries created before the year 2000.

```{r 3.1.2}
## extracts the entries before year 2000
## [WRITE CODE] ##
```

- Extract the entries where the `imports` value is larger than zero.

```{r 3.1.3}
## extracts the entries with imports > 0
## [WRITE CODE] ##
```

#### 3.2- Multiple conditions

- Extract the entries of the "EU countries" region in 1995. *Are there different ways to achieve this?*
- Assign the results to `ge_trade_1995`.

```{r 3.2.1}
## extract the entries of "EU countries" in 1995
ge_trade_1995 <- ge_trade %>% 
  ## [WRITE CODE] ##

## print result
ge_trade_1995
```

- Extract the entries in the year range [2000 to 2005]. 

**HINT:** You can use the `&` operator or `between`. 

```{r 3.2.2}
## extract the entries between [2000-2005]
## [WRITE CODE] ##
```

- Extract the entries with the following conditions:

    + year is 2015.
    + `exports_usd` value is larger than `imports_usd` value.

```{r 3.2.3}
## extract the entries of year 2015 where exports > imports
## [WRITE CODE] ##
```

### 4- summarize()

#### 4.1

using `ge_trade` data:

- filter the data to extract the entries corresponding to the year 2015.
- calculate the sum of `imports_usd` using `summarize`.

```{r 4.1.1}
## what's wrong with this?
ge_trade %>% 
  filter(year == 2015) %>% 
  summarize(imports = sum(imports_usd))
```

- Modify the previous code chunk to get the sum of `imports_usd`, excluding `NAs` 

```{r 4.1.2}
## modify the previous code chunk to exclude NAs
## [WRITE CODE] ##
```

### 5- group_by()

#### 5.1- Group by one variable

- Calculate the total value of imports and exports in each year.

**HINT:** Don't forget to remove `NAs`.

```{r 5.1.1}
## calculate imports and exports values for each year
ge_trade %>% 
  group_by(## [WRITE CODE] ##) %>% 
  summarize(imports = ## [WRITE CODE] ##),
            exports = ## [WRITE CODE] ##)
```

- Calculate the number of unique countries in each region in the dataset.

**HINT:** Consider using `n_distinct()`

```{r 5.1.2}
## calculate the number of unique countries in each region
ge_trade %>% 
  ## [WRITE CODE] ## %>% 
  summarize(countries = ## [WRITE CODE] ##)
```

#### 5.2- Group by multiple variables

- Calculate the total value of imports and exports for each `(region, year)` combination.

```{r 5.2.1}
## calculate the imports and exports values for each `(region, year)` combination 
## [WRITE CODE] ##
```

*Note that if you already have a grouped dataframe by `region`, you can can group it by another variable in a different step.*

```{r 5.2.2, eval=FALSE}
## return a grouped dataframe by region
ge_trade_region <- ge_trade %>% 
  group_by(region)  

## group by year in a different step
ge_trade_region %>% 
  group_by(year) %>% 
  summarize(imports = sum(imports_usd, na.rm = TRUE),
            exports = sum(exports_usd, na.rm = TRUE))
```

*Try switching `year` and `region` order and check the result.*

### 6- mutate()

#### 6.1- Add a new column

- Add a new column `exports_flag` with a flag to indicate whether `exports_usd` > `imports_usd`.

```{r 6.1.1}
## add exports_flag column
ge_trade %>% 
  mutate(## [WRITE CODE] ##)
```

- Add a new column `exports_over_imports` with the value of `exports` over `imports` in each row. 

```{r 6.1.2}
## add exports_over_imports column
ge_trade %>% 
  group_by(region, year) %>% 
  summarize(imports = sum(imports_usd, na.rm = TRUE),
            exports = sum(exports_usd, na.rm = TRUE)) %>% 
  ## [WRITE CODE] ##
```

#### 6.2- Change existing column

- Use mutate to modify `imports` and `exports` by dividing all the values by 1000.

```{r 6.2}
## modify the values on exports and imports
ge_trade %>% 
  group_by(region, year) %>% 
  summarize(imports = sum(imports_usd, na.rm = TRUE),
            exports = sum(exports_usd, na.rm = TRUE)) %>% 
  ## [WRITE CODE] ##
```

### 7- arrange()

#### 7.1

- Extract the values corresponding to year 2001.
- Calculate the total value of imports for each country. 
- Assign the result to `ge_imports`.

```{r 7.1.1}
## calculate imports per country in 2001
## [WRITE CODE] ##
```

- Arrange `ge_imports` based on the `imports` values.

```{r 7.1.2}
## arrange ge_imports based on `imports` values
ge_imports %>% 
  arrange(## [WRITE CODE] ##)
```

- Arrange `ge_imports` in a descending order based on the `imports` values.

```{r 7.1.3}
## arrange ge_imports in a descending order based on `imports` values
ge_imports %>% 
  ## [WRITE CODE] ##
```
